<!-- badges: start -->
[![R build status](https://github.com/ropenscilabs/rssr/workflows/R-CMD-check/badge.svg)](https://github.com/ropenscilabs/rssr/actions)
[![codecov](https://codecov.io/gh/ropenscilabs/rssr/branch/master/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/rssr)
[![Project Status:
Concept](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
<!-- badges: end -->


<!-- README.md is generated from README.Rmd. Please edit that file -->

# rssr

[**r**OpenSci](https://ropensci.org) **S**tatistical **S**software **R**eview.
This is an R package with functions to help those developing and reviewing
package submitted to rOpenSci's system for [Statistical Software
Review](https://ropenscilabs.github.io/statistical-software-review-book/index.html).
The functions are mostly intended to ease alignment and assessment of software
against the standards detailed in the [main project
book](https://ropenscilabs.github.io/statistical-software-review-book/index.html)
(from here on referred to as the "SSR Book").

This package can be installed with,

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("ropenscilabs/statistical-software-review")
```

and loaded for use with,
```{r lib-fakey, eval = FALSE}
library (rssr)
```
```{r lib, echo = FALSE, message = FALSE}
devtools::load_all (".", export_all = FALSE)
```

All functions of the package are prefixed with `rssr_`. The remainder of this
document is in two main sections. If you're developing a package for submission
to our peer review system, keep straight on reading. If you've been invited to
review a package, you may skip the following section and proceed straight to
the subsequent section. The general procedures for both developers and
reviewers are described at length in the [SSR
book](https://ropenscilabs.github.io/statistical-software-review-book/index.html),
with this `README` intended to provide supporting technical details.

## For Package Developers

People intending to develop packages for submission to our system for peer
reviewing statistical software will need to follow the following three general
steps. Note that, while the `rssr` package has a few functions which developers
may call directly to aid their submission process, most functionality of this
package is implemented via custom [`roxygen2`
"roclets"](https://roxygen2.r-lib.org). The third of the following steps
describes how to link your package with package in order to use these roclets.

1. Ensure the package successfully passes all
   [`autotest`](https://github.com/ropenscilabs/autotest) tests, including
   setting `test = FALSE` flags to switch off any particular tests. For
   details, please see the [package documentation for
   `autotest`](https://ropenscilabs.github.io/autotest/).
2. Decide which of our in-scope categories of statistical software best
   describe your package. The function
   [`rssr_available_categories()`](https://ropenscilabs.github.io/statistical-software-review/reference/rssr_available_categories.html)
   provides a list of currently developed categories for which standards
   have been developed, along with links to the online standards for each
   category:

    ```{r available}
    rssr_available_categories ()$title
    ```
    That function also returns links to the full descriptions of each category
    in the [main project
    book](https://ropenscilabs.github.io/statistical-software-review-book/index.html).
    Any software within one or more of these categories may be considered for
    review. 
3. Enable your package to use the `rssr` roclets by modifying the package's
   `DESCRIPTION` file so that the `Roxygen` line looks like this:
    ```{r roxygen, eval = FALSE}
    Roxygen: list(markdown = TRUE, roclets = c ("namespace", "rd", "rssr::rssr_roclet"))
    ```
    That will load the ["roclet"](https://roxygen2.r-lib.org) used by this
    package to insert standards within your actual code. Note that you do not
    need to add, import, or depend upon the `rssr` package anywhere else within
    the `DESCRIPTION` file.
4. Load the `rssr` package and generate lists of standards within your
   package's `/R` folder by running,
   [`rssr_standards_roxygen(category = c("<my-category-1>", "<my-category-2>"))`](https://ropenscilabs.github.io/statistical-software-review/reference/rssr_standards_roxygen.html).
   This will by default create a new file called by default
   `R/rssr_standards.R`, the first few lines of which will look like this:
    ```{r rssr_standards_roxygen, echo = FALSE, message = FALSE}
    f <- "./R/rssr-standards.R"
    if (!file.exists (f))
        rssr_standards_roxygen ()
    rso <- readLines (f)
    head (rso)
    ```
    The file will contain a list of all standards from your nominated
    categories. This file may be renamed, and the individual items moved to
    other locations in other files, but all nominated standards should remain
    somewhere in [`roxygen2`](https://roxygen2.r-lib.org) blocks somewhere in
    your source code.

    The `@rssrVerbose` line defines a variable which may be used to suppress
    output from the `rssr` roclet when updating package documentation (by
    setting to `FALSE`). After that comes the list of standards, each of which
    is prefixed by a [`roxygen2`](https://roxygen2.r-lib.org) tag, `@rssrTODO`.
    A package can only be submitted once all of these `TODO` items have been
    addressed via one of the options described in the following two items.
5. A standard may be addressed by moving the item in the `rssr_standards.R`
   file (or wherever you've chosen to list these within your own package) to
   one or more places in your code where these standards have been addressed.
   In doing so, the [`roxygen2`](https://roxygen2.r-lib.org) tag should be
   changed from `@rssrTODO` to `@rssr`, and the text which initially lists the
   actual standard should be changed to provide a brief description of how that
   standard has been met. Tags for one particular standard may be repeated in
   multiple places within your code, and we encourage locating an `@rssr` tag
   reference to a particular standard at all locations which directly address
   that standard.
6. Alternatively, any standards which you consider not applicable to your
   software may remain listed in the main `rssr_standards.R` document, with
   their tag changed from `@rssrTODO` to `@rssrNA`, and the description of the
   standard removed and replaced by an explanation of why you consider that
   standard not applicable to your software.
7. Each time you run 
    [`devtools::document()`](https://devtools.r-lib.org/reference/document.html)
    or the equivalent
    [`roxygen2::roxygenise()`](https://roxygen2.r-lib.org/reference/roxygenize.html),
    the roclet will scan your package's documentation for the state of
    standards, and will generate a summary of the result on your screen.

Note that individual standards may be moved to, and addressed in, any location
including the directories `R/`, `src/`, or `tests/`. The [`roxygen2`
roclet](https://roxygen2.r-lib.org) associated with this package is able to
parse the various `@rssr` tags in all of these locations. Note, however, that
tags in `src/` directories are only able to be parsed form C++ files compiled
with [`Rcpp`](https://cran.r-project.org/package=Rcpp). Equivalent parsing of
C++ packages compiled with [`cpp11`](https://cpp11.r-lib.org) should hopefully
[soon be possible](https://github.com/r-lib/cpp11/issues/147).
