<!-- badges: start -->
[![R build status](https://github.com/ropenscilabs/rrr/workflows/R-CMD-check/badge.svg)](https://github.com/ropenscilabs/rrr/actions)
[![codecov](https://codecov.io/gh/ropenscilabs/rrr/branch/master/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/rrr)
[![Project Status:
Concept](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
<!-- badges: end -->


<!-- README.md is generated from README.Rmd. Please edit that file -->

# [**r**OpenSci](https://ropensci.org) **r**eview **r**oclets (rrr)j

This package contains roclets for use with [rOpenSci](https://ropensci.org)'s
software review systems. At present, the package only contains roclets and
associated functions to help those developing and reviewing packages submitted
to rOpenSci's system for [Statistical Software
Review](https://ropenscilabs.github.io/statistical-software-review-book/index.html).
The functions are mostly intended to ease alignment and assessment of software
against the standards detailed in the [main project
book](https://ropenscilabs.github.io/statistical-software-review-book/index.html)
(from here on referred to as the "SSR Book").

This package can be installed with,

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("ropenscilabs/rrr")
```

and loaded for use with,
```{r lib-fakey, eval = FALSE}
library (rrr)
```
```{r lib, echo = FALSE, message = FALSE}
devtools::load_all (".", export_all = FALSE)
```

Both this `README`, and the main package vignette, describe the functionality
of the package in the specific context of the statistical software review
project. Both the roclet and all functions intended for use in this context are
prefixed with `rrr_stats_`. The remainder of this document is in two main
sections. If you're developing a statistics package for submission to our peer
review system, keep straight on reading. If you've been invited to review a
package, you may skip the following section and just read the subsequent
section. The general procedures for both developers and reviewers are described
at length in the [SSR
book](https://ropenscilabs.github.io/statistical-software-review-book/index.html),
with this `README` intended to provide supporting technical details.

## For Package Developers

People intending to develop packages for submission to our system for peer
reviewing statistical software will need to follow the following general steps.
Note that, while the `rrr` package has a few functions which developers may
call directly to aid their submission process, most functionality of this
package is implemented via custom [`roxygen2`
"roclets"](https://roxygen2.r-lib.org). The third of the following steps
describes how to link your package with `rrr` in order to use these roclets.

1. Ensure that your package successfully passes all
   [`autotest`](https://github.com/ropenscilabs/autotest) tests, potentially
   including setting `test = FALSE` flags to switch off any tests you consider
   not to be applicable to your package. For details, please see the [package
   documentation for `autotest`](https://ropenscilabs.github.io/autotest/).
2. Decide which of our in-scope categories of statistical software best
   describe your package. The function
   [`rrr_stats_categories()`](https://ropenscilabs.github.io/rrr/reference/rrr_stats_categories.html)
   provides a list of currently developed categories for which standards
   have been developed, along with links to the online standards for each
   category:

    ```{r available}
    rrr_stats_categories ()$title
    ```
    That function also returns links to the full descriptions of each category
    in the [main project
    book](https://ropenscilabs.github.io/statistical-software-review-book/index.html).
    Any software within one or more of these categories may be considered for
    review. 
3. Enable your package to use the `rrr_stats` roclets by modifying the
   package's `DESCRIPTION` file so that the `Roxygen` line looks like this:
    ```{r roxygen, eval = FALSE}
    Roxygen: list(markdown = TRUE, roclets = c ("namespace", "rd", "rrr::rrr_stats_roclet"))
    ```
    That will load the ["roclet"](https://roxygen2.r-lib.org) used by this
    package to process the documentation of statistical standards within your
    actual code. Note that you do not need to add, import, or depend upon the
    `rrr` package anywhere else within the `DESCRIPTION` file or your actual
    package.
4. Load the `rrr` package and generate lists of standards within your
   package's `/R` folder by running,
   [`rrr_stats_roxygen(category = c("<my-category-1>", "<my-category-2>"))`](https://ropenscilabs.github.io/rrr/reference/rrr_stats_roxygen.html).
   This will by default create a new file called by default
   `R/rrr_stats_standards.R`, the first few lines of which will look like this:
    ```{r rrr_stats_roxygen, echo = FALSE, message = FALSE}
    f <- "./R/rrr-stats-standards.R"
    if (!file.exists (f))
        rrr_stats_roxygen ()
    rso <- readLines (f)
    head (rso)
    ```
    The file will contain a list of all standards from your nominated
    categories. This file may be renamed, and the individual items moved to
    other locations in other files, but all nominated standards should remain
    in [`roxygen2`](https://roxygen2.r-lib.org) blocks somewhere in your source
    code.

    The `@rrrstatsVerbose` line defines a variable which may be used to
    suppress output from the `rrrstats` roclet when updating package
    documentation (by setting to `FALSE`). After that comes the list of
    standards, each of which is prefixed by
    a [`roxygen2`](https://roxygen2.r-lib.org) tag, `@rrrstatsTODO`. A package
    can only be submitted once all of these `TODO` items have been addressed
    via one of the options described in the following two items.
5. A standard may be addressed by moving the item in the
   `rrr-stats-standards.R` file (or wherever you've chosen to list these within
   your own package) to one or more places in your code where these standards
   have been addressed. In doing so, the
   [`roxygen2`](https://roxygen2.r-lib.org) tag should be changed from
   `@rrrstatsTODO` to `@rrrstats`, and the text which initially lists the
   actual standard should be changed to provide a brief description of how that
   standard has been met. Tags for one particular standard may be repeated in
   multiple places within your code, and we encourage locating multiple
   `@rrrstats` tags which refer to a particular standard at all locations which
   directly address that standard.
6. Alternatively, any standards which you consider not applicable to your
   software may remain listed in the templated section of the main
   `rrr-stats-standards.R` document (or any alternative location), with their
   tag changed from `@rrrstatsTODO` to `@rrrstatsNA`, and the description of
   the standard removed and replaced by an explanation of why you consider that
   standard not to be applicable to your software. These `@rrrstatsNA` tags
   should be collected together within a single `roxygen2` block with a title
   of `NA_standards`, as provided in the initial template generated by the
   [`rrr_stats_roxygen()`](https://ropenscilabs.github.io/rrr/reference/rrr_stats_roxygen.html)
   function. Any non-applicable standards can then just be moved into this
   block, with their `@rrrstatsTODO` tags changed to `@rrrstatsNA` 
7. Each time you run 
    [`devtools::document()`](https://devtools.r-lib.org/reference/document.html)
    or the equivalent
    [`roxygen2::roxygenise()`](https://roxygen2.r-lib.org/reference/roxygenize.html),
    the roclet will scan your package's documentation for the state of
    standards, and will generate a summary of the result on your screen.

To help developers understand how to use these roclets, this package includes a
function,
[`rrr_stats_pkg_skeleton()`](https://ropenscilabs.github.io/rrr/reference/rrr_stats_pkg_skeleton.html),
which will generate a skeleton of a package with several
[`roxygen2`](https://roxygen2.r-lib.org) tags inserted throughout the code.
This function returns the directory where the skeleton package has been
created, so running the following two lines will illustrate the roclets in
action:

```{r demo, eval = FALSE}
d <- rrr_stats_pkg_skeleton ()
roxygen2::roxygenise (d)
```

Note that the skeleton package also includes C++ code in a `src/` directory, so
will be compiled the first time your run
[`roxygensise()`](https://roxygen2.r-lib.org/reference/roxygenize.html).
Running a second time will generate cleaner output from the `rrrstats` roclets
only. The tags included in 
[`roxygen2`](https://roxygen2.r-lib.org/) blocks in this skeleton package may
be modified, moved, copied, and changed in any way you like to help you
understand how the roclets work. Simply play around with the 
[`roxygen2`](https://roxygen2.r-lib.org/) lines and run 
[`roxygensise()`](https://roxygen2.r-lib.org/reference/roxygenize.html) each
time to see the effect. Individual standards may be moved to, and addressed in,
any location including the directories `R/`, `src/`, or `tests/`. The
`rrr_stats` roclet associated with this package is able to parse the various
`@rrrstats` tags in all of these locations.
